Predicting The Outcome T1 and Gen G Games Based on Leads

### 2024-04-19 paper/MatchPrediction.qmd
renamed some files

author: 
  - Rahul Gopeesingh
thanks: "Code and data are available at:https://github.com/Rahul-Uoft/Match-Prediciton.git"
date: today
date-format: long
abstract: "T1 and Gen G are two of the best League of Legends teams in the world with a longstanding rivalry. In a game of league of legends, the early focus of the game is not the primary objective, but rather, gain advantages so that it is easier to achieve the primary objective. Some teams are really good at throwing away the advantages they did so well to accrue. This paper investigates how good T1 and Gen G are at converting their leads into a win. The model predicts that Gen G is better at playing with leads, and are more likely to convert to a win, but are much worse playing without them. This information can be used by coaching staff to gain knowledge about the players as well as other teams to devise strategies to play against them."
format: pdf
number-sections: true
bibliography: references.bib
toc: TRUE
---

@@

In this paper we predict the outcome of a given game of league of legends played by players of the pro teams T1 and Generation Gaming(Gen G) based on their creep score(CS) at 10 minutes into the game and the highest level lead they have against their opponent.

@@

The two gaming organizations, T1 and Gen G, have been battling at the top level for the last few years. They are both Korean teams, with Korea being arguably the strongest region in the world. At the time of writing this paper, T1 are the current world champions, and have some of the highest caliber of players. They were the favourites to win the 2023 world championship and they are home to the most renowned and respected league of legends pro player of all time, Faker. However there is a caveat to their apparent dominance. Domestically, they struggle to beat Gen G. Many fans coin Gen G as 'T1's Kryptonite.' In the 2024 Korean Championship, Gen G beat T1 in a very close best of 5 that ended 3-2. Many fans also believe that T1 only won the world championship because Gen G got knocked out by a Chinese team.

@@

The data used in this paper was downloaded directly fromt the Riot Games API [@riotAPI], the company that owns league of legends. This is available to everyone for free(although an account must be created.) When a game of league of legends is played online, a large amount of data is recorded and stored in their database. These can then be accessed using their API, which allows this data to be queried. The data used in this paper comes from games played by each of the players on the pro teams, and were within the last 2 years. This allows it to be more indicative of their current performance, rather than their performance from 2016. It is taken from their solo queue games, which is essentially the regular games that anyone can play online by themselves and not the professional games in which all of the team's players team up and play against another full team. (As the name suggests, solo queue means you are queueing for a game alone.)

@@

There are 3 variables that we are concerned with, the player's creep score(cs) at 10 minutes into the game, the player's highest level lead against his opponent, and the outcome of the game. The implications of these are discussed thoroughly in the discussion but the overarching belief is that the stronger you are, the easier it is to win the game. The higher the cs at 10 minutes is,  the stronger you are as you have more gold to buy items. The same goes for the level disparity between you and your opponent. However, being stronger does not guarentee a win. Some players may be better at converting their leads into victories and therefore we analyze one of the richest rivalries in league of legends history, T1 vs Generation Gaming.

@@

Another advantage of using a logistic regression model is that it does not require the independence of trials. This is great as burnout is a very real thing when playing league of legends. Players frequently go on both win streaks and loss streaks which would indicate that your recent games can impact your current one. This could easily make some of these trials dependent on some of each other but this won't affect the usability of the logistic regression model. We use the default priors in this model as default priors aren't very informative and therefore let the data about the games drive the model without much interference.

@@

A major weakness of this analysis is that the the games under consideration were the solo queue games, with random teammates of a similar level. This version of league of legends has many differences to the professional play where team coordination is much more important than any individual player's skill level. These games were used as they are readily avaailable and generally speaking, the best professional players do really well in solo queue. Their solo queue characteristics are also apparent in their professional games. An example of this would Chovy, Gen G's mid laner. He is renowned for his extremely high cs and this shows up in both solo queue and professional matches. In this way, solo queue data are good approximations for analyzing certain aspects of player performance however the professional game data would be much more relevant for predicting professional games. Another weakness is the relatively small dataset. These players play thousands of games every year and we can analyze all of them to get a better understanding of how they play without cyclical shifts such as meta changes.

>>>>

author: 
  - Rahul Gopeesingh
thanks: "Code and data are available at:https://github.com/Rahul-Uoft/Match-Prediciton.git"
date: today
date-format: long
abstract: "T1 and Gen G are two of the best League of Legends teams in the world with a longstanding rivalry. In a game of League of Legends, the early focus of the game is not the primary objective, but rather, gain advantages so that it is easier to achieve the primary objective. Some teams are really good at throwing away the advantages they did so well to accrue. This paper investigates how good T1 and Gen G are at converting their leads into a win. The model predicts that Gen G is better at playing with leads, and are more likely to convert to a win, but are much worse playing without them. This information can be used by coaching staff to gain knowledge about the players as well as other teams to devise strategies to play against them."
format: pdf
number-sections: true
bibliography: references.bib
toc: TRUE
---

@@

In this paper we predict the outcome of a given game of League of Legends played by players of the pro teams T1 and Generation Gaming(Gen G) based on their creep score(CS) at 10 minutes into the game and the highest level lead they have against their opponent.

@@

The two gaming organizations, T1 and Gen G, have been battling at the top level for the last few years. They are both Korean teams, with Korea being arguably the strongest region in the world. At the time of writing this paper, T1 are the current world champions, and have some of the highest caliber of players. They were the favourites to win the 2023 world championship and they are home to the most renowned and respected League of Legends pro player of all time, Faker. However there is a caveat to their apparent dominance. Domestically, they struggle to beat Gen G. Many fans coin Gen G as 'T1's Kryptonite.' In the 2024 Korean Championship, Gen G beat T1 in a very close best of 5 that ended 3-2. Many fans also believe that T1 only won the world championship because Gen G got knocked out by a Chinese team.

@@

The data used in this paper was downloaded directly fromt the Riot Games API [@riotAPI], the company that owns League of Legends. This is available to everyone for free(although an account must be created.) When a game of League of Legends is played online, a large amount of data is recorded and stored in their database. These can then be accessed using their API, which allows this data to be queried. The data used in this paper comes from games played by each of the players on the pro teams, and were within the last 2 years. This allows it to be more indicative of their current performance, rather than their performance from 2016. It is taken from their solo queue games, which is essentially the regular games that anyone can play online by themselves and not the professional games in which all of the team's players team up and play against another full team. (As the name suggests, solo queue means you are queueing for a game alone.)

@@

There are 3 variables that we are concerned with, the player's creep score(cs) at 10 minutes into the game, the player's highest level lead against his opponent, and the outcome of the game. The implications of these are discussed thoroughly in the discussion but the overarching belief is that the stronger you are, the easier it is to win the game. The higher the cs at 10 minutes is,  the stronger you are as you have more gold to buy items. The same goes for the level disparity between you and your opponent. However, being stronger does not guarentee a win. Some players may be better at converting their leads into victories and therefore we analyze one of the richest rivalries in League of Legends history, T1 vs Generation Gaming.

@@

Another advantage of using a logistic regression model is that it does not require the independence of trials. This is great as burnout is a very real thing when playing League of Legends. Players frequently go on both win streaks and loss streaks which would indicate that your recent games can impact your current one. This could easily make some of these trials dependent on some of each other but this won't affect the usability of the logistic regression model. We use the default priors in this model as default priors aren't very informative and therefore let the data about the games drive the model without much interference.

@@

A major weakness of this analysis is that the the games under consideration were the solo queue games, with random teammates of a similar level. This version of League of Legends has many differences to the professional play where team coordination is much more important than any individual player's skill level. These games were used as they are readily avaailable and generally speaking, the best professional players do really well in solo queue. Their solo queue characteristics are also apparent in their professional games. An example of this would Chovy, Gen G's mid laner. He is renowned for his extremely high cs and this shows up in both solo queue and professional matches. In this way, solo queue data are good approximations for analyzing certain aspects of player performance however the professional game data would be much more relevant for predicting professional games. Another weakness is the relatively small dataset. These players play thousands of games every year and we can analyze all of them to get a better understanding of how they play without cyclical shifts such as meta changes.
